---
title: "Moorea Coral"
author: 
    name: "Monique Hernandez"
    affiliation: University of California, Santa Barbara
date: "Augustt 18, 2026"
format:
  html:
    toc: true
execute:
  message: false
  warning: false
---

[click here for github link](https://github.com/moniqueallyson/eds221day6) 

# Read Data
## Moorea Coral Data Frame
```{r, message=FALSE}

library(tidyverse)

moorea_coral <- read.csv("data/moorea_coral.csv", na = c("", "NA", "ND")) # read moorea_coral.csv and remove missing data

glimpse(moorea_coral) # preview dataset

```

# Visualize Data
## Exericse 1 & 2. Percent Cover by Habitat, Site, and Taxa
```{r}


moorea_plot <- moorea_coral |> # re-assigning coral data to plot
    filter(Site == "LTER_1") |> # filter by Site
    filter(Date == "2017-04") |> # filter by Date
    filter(Taxonomy_Substrate_or_Functional_Group == "Acropora" | Taxonomy_Substrate_or_Functional_Group == "Pocillopora") |> # filter by Taxonomy
    rename(Taxa = Taxonomy_Substrate_or_Functional_Group) # rename taxnomy


ggplot(
  data = moorea_plot, # insert data for ggplot
  mapping = aes(
    x = Habitat, # x-axis
    y = Percent_Cover, # y-axis
    fill = Taxa #
    )
) +
  geom_boxplot() + 
  scale_y_continuous(
    name = "Percent Cover"
  )


moorea_plot |> # data frame
    mutate(Percent_Cover = Percent_Cover / 100) |> # nutate percent cover by dividing by 100
        summarize( # summarize
    avg_size = mean(Percent_Cover, na.rm = TRUE), # percent cover mean
    stan_dev = sd(Percent_Cover, na.rm = TRUE), # percent cover standard deviation
    .by = c(Habitat, Depth, Taxa)) # .by 


```



## Exercise 3. Findings 
First, the library(tidyverse) and the csv file containing the Moorea Coral dataset were downloaded and read. Then, a boxplot is visualized by utilizing the aesthetic features to give the x-and-y axis, which are categorized respectively by habitat and percent cover for Acropora and Pocillopora. Lastly, the mean and standard deviation are shown grouped by habitat, depth, and taxa. 